### 2020-12-11 ch01_introduction/introduction.tex
Added climate feedback figure and adapted margins
Added overview of climate feedbacks
Started section on techniques for reducing uncertainties in climate model projections

@@

\section{Structure of the thesis}
\label{sec:01:structure}

Parts of this thesis are published in multiple peer-reviewed publications (two
first-author studies and six co-author studies). If applicable, this is clearly
stated at the beginning of each chapter. \Cref{ch:02:scientific_background}
introduces the scientific background for this thesis. This includes relevant
literature that is used as a baseline for this thesis. \Cref{ch:03:esmvaltool}
gives an overview over the contributions made to the \ac{ESMValTool}, an
open-source software for the analysis of \acp{ESM}. These contributions helped
improving the routine evaluation of \acp{ESM} which is useful for the whole
scientific community and lead to co-authorship in four peer-reviewed studies
\autocite{Righi2020, Eyring2020, Lauer2020, Weigel2020}.
\Cref{ch:04:assessment_climate_metrics} covers the assessment of
policy-relevant climate metrics like the \ac{ECS} and the \ac{TCR} in the
latest generation of \acp{ESM}. This work is already published in two
scientific publications \autocite{Bock2020, Meehl2020}. Since the \ac{ECS} and
\ac{TCR} are considerably higher in this new climate model generation,
\cref{ch:05:paper_ecs} describes the assessment of emergent constraints (a
technique to reduce uncertainties in climate model projections, see
\vref{sec:02:techniques}) on the \ac{ECS} for these \acp{ESM}. The contents of
this chapter are published in \emph{Earth System Dynamics}
\autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
reduce climate model uncertainties based on \ac{ML}. As an example, the method
is applied to the photosynthesis rate at the end of the \nth{21} century, which
is already published in the \emph{Journal of Geophysical Research:
  Biogeosciences} \autocite{Schlund2020}. Finally, \cref{ch:07:summary_outlook}
provides a summary of the results of this thesis and gives an outlook of
possible future works.

>>>>

- definition of radiative forcing

@@

\section{Structure of the Thesis}
\label{sec:01:structure}

Parts of this thesis are published in multiple peer-reviewed publications (two
first-author studies and six co-author studies). If applicable, this is clearly
stated at the beginning of each chapter. \Cref{ch:02:scientific_background}
introduces the scientific background for this thesis. This includes relevant
literature that is used as a baseline for this thesis. \Cref{ch:03:esmvaltool}
gives an overview over the contributions made to the \ac{ESMValTool}, an
open-source software for the analysis of \acp{ESM}. These contributions helped
improving the routine evaluation of \acp{ESM} which is useful for the whole
scientific community and lead to co-authorship in four peer-reviewed studies
\autocite{Righi2020, Eyring2020, Lauer2020, Weigel2020}.
\Cref{ch:04:assessment_climate_metrics} covers the assessment of
policy-relevant climate metrics like the \ac{ECS} and the \ac{TCR} in the
latest generation of \acp{ESM}. This work is already published in two
scientific publications \autocite{Bock2020, Meehl2020}. Since the \ac{ECS} and
\ac{TCR} are considerably higher in this new climate model generation,
\cref{ch:05:paper_ecs} describes the assessment of emergent constraints (a
technique to reduce uncertainties in multi-model climate projections, see
\vref{subsec:02:emergent_constraints}) on the \ac{ECS} for these \acp{ESM}. The
contents of this chapter are published in \emph{Earth System Dynamics}
\autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
reduce uncertainties in multi-model projections of the future climate based on
\ac{ML}. As an example, the method is applied to the photosynthesis rate at the
end of the \nth{21} century, which is already published in the \emph{Journal of
  Geophysical Research: Biogeosciences} \autocite{Schlund2020}. Finally,
\cref{ch:07:summary_outlook} provides a summary of the results of this thesis
and gives an outlook of possible future works.

### 2020-12-11 ch02_scientific_background/scientific_background.tex
Started section on techniques for reducing uncertainties in climate model projections

Similar to \ac{ECS}, \ac{TCR} can also be defined in terms of an external
forcing and climate feedbacks. For this, the energy balance
\cref{eq:02:N_vs_delta_t} can be slightly adjusted. Since over \pct{90} of the
excess energy introduced into the system by the radiative forcing $F$ is taken
up by the ocean due to its large heat capacity, $N$ can be taken equal to the
global ocean heat uptake \autocite{Knutti2017}. In experiments with a steadily
increasing radiative forcing, which is the case for the \onepctcotwo{}
simulation, this ocean heat uptake can be approximated with
\begin{equation}
  N = \kappa \cdot \Delta T,
  \label{eq:02:ocean_heat_uptake}
\end{equation}
where $\kappa$ is the ocean heat uptake efficiency \autocite{Gregory2008}.
$\kappa > 0$, since there is a net energy flux into the climate system ($N >
0$) due to the external forcing $F$. This approximation becomes less accurate
as the deeper ocean warms up and cannot be applied to simulations with a steady
state climate change ($N \to 0$) where the external forcings tends stabilize
\autocite{Gregory2009}. By applying the definition of \ac{TCR} (transient
\ac{GSAT} change $\Delta T$ at the time of \ac{CO2} doubling) and combining
\cref{eq:02:N_vs_delta_t,eq:02:ocean_heat_uptake}, \ac{TCR} can be estimated as
\begin{equation}
  \text{TCR} = \frac{F_\text{2x}}{\kappa - \lambda},
  \label{eq:02:tcr}
\end{equation}
where $F_\text{2x}$ is the radiative forcing induced by a doubling of the
atmospheric \ac{CO2} concentration.

This equation can be used to derive a relationship between \ac{TCR} and
\ac{ECS}. Writing $\text{ECS} = -F_\text{2x} / \lambda$ and assuming path
independence of the forcing (\ie{} the resulting radiative forcing from the
\onepctcotwo{} and \nxcotwo{2} runs are comparable) gives
\begin{equation}
  \text{TCR} = \frac{F_\text{2x} \cdot \text{ECS}}{F_\text{2x} + \kappa \cdot
    \text{ECS}},
  \label{eq:02:tcr_vs_ecs}
\end{equation}
which demonstrates the non-linear connection between \ac{TCR} and \ac{ECS} \autocite{Gregory2008, Nijsse2020}.

@@

\section{Techniques to Reduce Uncertainties in Climate Model Projections}
\label{sec:02:techniques}

- Process-based weighting (Knutti + MDER)

- emergent constraints

@@

>>>>

Similar to \ac{ECS}, \ac{TCR} can also be defined in terms of an external
forcing and climate feedbacks. For this, the energy balance
\cref{eq:02:N_vs_delta_t} can be slightly adjusted. Since over \pct{90} of the
excess energy introduced into the climate system by the radiative forcing $F$
is taken up by the ocean due to its large heat capacity, $N$ can be taken equal
to the global ocean heat uptake \autocite{Knutti2017}. In experiments with a
steadily increasing radiative forcing, which is the case for the \onepctcotwo{}
simulation, this ocean heat uptake can be approximated with
\begin{equation}
  N = \kappa \cdot \Delta T,
  \label{eq:02:ocean_heat_uptake}
\end{equation}
where $\kappa$ is the ocean heat uptake efficiency \autocite{Gregory2008}.
Since there is a net energy flux into the climate system ($N > 0$) due to the
external forcing $F$, $\kappa$ is positive. This approximation becomes less
accurate as the deeper ocean warms up and cannot be applied to simulations with
a steady state climate change ($N \to 0$) where the external forcing tends
stabilize \autocite{Gregory2009}. By applying the definition of \ac{TCR}
(transient \ac{GSAT} change $\Delta T$ at the time of \ac{CO2} doubling) and
combining \cref{eq:02:N_vs_delta_t,eq:02:ocean_heat_uptake}, \ac{TCR} can be
estimated as
\begin{equation}
  \text{TCR} = \frac{F_\text{2x}}{\kappa - \lambda},
  \label{eq:02:tcr}
\end{equation}
where $F_\text{2x}$ is the radiative forcing induced by a doubling of the
atmospheric \ac{CO2} concentration.

This equation can be used to derive a relationship between \ac{TCR} and
\ac{ECS}. Writing $\text{ECS} = -F_\text{2x} / \lambda$ and assuming path
independence of the forcing (\ie{} the resulting radiative forcing from the
\onepctcotwo{} and \nxcotwo{2} runs are comparable) gives
\begin{equation}
  \text{TCR} = \frac{F_\text{2x} \cdot \text{ECS}}{F_\text{2x} + \kappa \cdot
    \text{ECS}},
  \label{eq:02:tcr_vs_ecs}
\end{equation}
which demonstrates the non-linear connection between \ac{TCR} and \ac{ECS}
\autocite{Gregory2008, Nijsse2020}. Since $\lambda < 0$ and $\kappa > 0$,
\cref{eq:02:tcr} implies that the equilibrium response \ac{ECS} is (as
expected) larger than transient response \ac{TCR}, \ie{} $\text{ECS} >
\text{TCR}$.

@@

\section{Techniques to Reduce Uncertainties in Multi-Model Climate Projections}
\label{sec:02:techniques_to_reduce_uncertainties}

As shown in \cref{subsec:02:source_of_uncertainties}, projections of the future
climate are always associated with uncertainties. In the context of this
thesis, the most relevant source for these is the climate response uncertainty.
It originates from the different responses of the different climate models in a
multi-model ensemble to a given forcing and affects all multi-model projections
of the future climate (\eg{} the climate sensitivity or the future development
of the global carbon cycle). This section introduces three state-of-the-art
techniques to reduce uncertainties in multi-model climate model projections
with observations. The presented methods form the baseline for the work
presented in this thesis.

\subsection{Emergent Constraints}
\label{subsec:02:emergent_constraints}

- equations for emergent constraints, see Schlund ESD, following Cox 2013

- Flat prior, see Schlund ESD

@@

\subsection{Performance- and Interdependence-based Weighting of Climate Models}
\label{subsec:02:model_weighting}

- Knutti, Sanderson

- recent studies: Brunner (T), Paul Young (Ozone), Lorenz?

- Lorenz (2018): reasoning for non-equal weighting, Collins 2013 as bad example

- Basic assumption: Models good in past -> good in future (PERFORMANCE)

- Models share code (INTERDEPENDENCE)

- Review by Abramowitz

- Theory: Knutt, p.1910

\subsection{\acl{MDER}}
\label{subsec:02:mder}

- some equations for MDER

- Karpechko, Wenzel, Senftleben

- basic assumption (in contrast to weighting), see Schlund JGR 2020
